---
title: "MSBA7002_Tutorial_2"
author: "Yutao DENG"
date: "2023-11-10"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
    number_sections: yes
    latex_engine: xelatex
  word_document:
    toc: yes
    toc_depth: '4'
  html_document:
    code_folding: show
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
---

\fontsize{12}{22}

\newpage

```{r setup, include=FALSE}
# knitr::opts_chunk$set(results='hide', tidy = T)
```


# Formating Markdown

## Equations 

$$
\begin{aligned}
L(\beta|X) &= \prod_{i=1}^n p_i^{a}(1-p_i)^{b}\\
log(L(\beta|X) )&=\sum_{i=1}^n (a * log(p_i) + b * log(1-p_i))
\end{aligned}
$$


$$
X =
\left\{\begin{array}{rl}
                  {1} & \text{true} \\
                  {0} & \text{false} \\
                \end{array}
              \right.
$$

## Itemize and Enumerate

For PDF output:
\begin{itemize}
  \item 1111
  \item 222
  \item 33
\end{itemize}


\begin{enumerate}
  \item AAA
  \item BB
  \item C
\end{enumerate}


For HTML output: (remember to leave an empty line before the list)

* Item A
* Item B
  + Item B1

1. Item 1
2. Item 2
3. Item 3
4. Item 4
    + Item 3a
    + Item 3b



# Package Management 

## Install packages in R
```{r}
# Install pacman package
# install.packages("pacman")

# Install package from url or file
# packageurl <- "https://cran.r-project.org/src/contrib/Archive/DiscriMiner/DiscriMiner_0.1-29.tar.gz"
# install.packages(packageurl, repos=NULL, type="source")

# Load pacman package
library("pacman")

# Detach pacman package
detach("package:pacman")
```


## Using pacman
```{r}
# package list for today: pROC,caret,dplyr,e1071

#Fisrt import the pacman package
# library("pacman")
# p_load(pROC,caret,dplyr,e1071)
# p_unload(pROC,caret,dplyr,e1071)

#Without loading the pacman package
# detach("package:pacman")
pacman::p_load(ggcorrplot, corrplot, car, nnet, GGally,kableExtra,pROC,caret,dplyr,e1071)

```
# Data manipulation

## Using dplyr
```{r}
# Load the data of smartphone
sp <- read.csv("smartphone.dat")

# colnames(sp) # without dplyr
sp %>% 
  colnames()
```

```{r}
# sum(is.na(sp)) # without dplyr
sp %>% 
  is.na() %>% 
  sum()
```

```{r}
# select from sp data where Group is Med
sp %>% 
  filter(Group=='Med')

# select from sp data where Group is Med or Low
sp %>% filter(Group %in% c('Med', 'Low'))
```

```{r}
# select some column by name 
sp %>% 
  dplyr::select('Age','Group')

# select some column by index 
sp %>% 
  dplyr::select(c(2,4))

# use select and filter at the same time
sp %>% 
  dplyr::select(c(1,2,4)) %>% 
  dplyr::filter(Group %in% c('Med', 'Low'))
```

```{r}
# gourp by Group
grp_sp <- sp %>% group_by(Group)
sp
grp_sp
grp_sp %>% summarise(sum(Income))
```

# Formating Results

## Using kableExtra
```{r}
# Basic HTML table
kbl(head(sp))

# bootstrap theme
sp %>%
  head() %>%
  kbl() %>%
  kable_styling()

# add caption
sp %>%
  head() %>%
  kbl(caption = "Smartphone data table") %>%
  kable_styling()


# Try other themes by replacing the last line
# kable_paper, kable_classic, kable_classic_2, kable_minimal, kable_material and kable_material_dark
sp %>%
  head() %>%
  kbl(caption = "Smartphone data table", centering = T, align = c('r', 'r', 'r')) %>%
  kable_classic(full_width = T, html_font = "Cambria")

# Add more details
sp %>%
  head() %>%
  kbl(caption = "Smartphone data table", centering = T, align = c('r', 'r', 'r')) %>%
  kable_styling() %>%
  add_header_above(c(" AIR " = 3, "Categorial" = 1))%>%
  pack_rows(index = c("1-3" = 3, "4-6" = 3))

# For pdf output
# , format="latex"
sp %>%
  head() %>%
  kbl(booktabs = TRUE, format="latex", escape = F, align = c('r', 'r', 'r'), 
      col.names = c("","$\\Pi$", "$\\hat{a}$", "$R^2$")) %>%
  kable_styling(latex_options = "striped", full_width = TRUE, font_size = 12)
```


## Figures

### Figures for EDA

```{r, warning=F, message=F}
# Use the iris data in GGally
iris%>%colnames()

# # scatter plots
# # Use geom_point
# iris%>%
#   filter(Species=="setosa")%>%
#   ggplot(aes(x = Petal.Length, y = Petal.Width)) +
#   geom_point()

# iris%>%
#   filter(Species=="setosa")%>%
#   dim()
# 
# # Add transparency to show all points
# iris%>%
#   filter(Species=="setosa")%>%
#   ggplot(aes(x = Petal.Length, y = Petal.Width)) +
#   geom_point(alpha = 0.3, size = 2.0)
# 
# # Also can use geom_jitter to add random shift
# iris%>%
#   filter(Species=="setosa")%>%
#   ggplot(aes(x = Petal.Length, y = Petal.Width)) +
#   geom_jitter(width=0.05, height=0.05)
# 
# 
# # Set colors for each group
# iris%>%
#   ggplot(aes(x = Petal.Length,
#              y = Petal.Width,
#              color = Species)
#          ) +
#   geom_jitter(width=0.05, height=0.05)
# 
# # Set different size by the sepal.length
# iris%>%
#   ggplot(aes(x = Petal.Length,
#              y = Petal.Width,
#              color = Species,
#              size = Sepal.Length)
#          ) +
#   geom_jitter(width=0.05, height=0.05, alpha = 0.4)

```



```{r, warning=F, message=F}
# # ggpairs: generalized pairs plot
# ggpairs(iris)
# 
# # select columns and color by group
# iris %>%
#   ggpairs(columns = 1:4,      # Columns
#         aes(color = Species,  # Color by group (cat. variable)
#             alpha = 0.5))     # Transparency
# 
# # Figure for correlation matrix
# # Use ggcorrplot
# iris[,1:4]%>%
#   cor()%>%
#   ggcorrplot(hc.order=TRUE)
# 
# # Use corrplot
# iris[,1:4]%>%
#   cor()%>%
#   corrplot()

```



### Figures for linear models
```{r}
# Recall the linear regression in last tutorial using smartphone data
# lm_fit <- lm(Rating ~ Income, data = sp)

# sp%>%
#   ggplot(aes(x = Rating, y = Income, color = Group)) +
#   geom_point() +
#   geom_smooth(method = "lm", formula = "y~x")
```


# Logistic regression
```{r}
# Use the titanic data
tit <- read.csv("titanicpassengers-bbm.dat")
str(tit)
tit$Passenger.Class <- factor(tit$Passenger.Class)
tit$Port <- factor(tit$Port, levels = c("C","Q","S"))
tit$Survived <- factor(tit$Survived)
# glm() generalized linear model 
## it requires y as a factor varaible
```

```{r}
# check the missing value
apply(tit, 2, function(x) sum(is.na(x)))

# Remove the rolls with missing values
# since we want to keep the age column
tit.comp <- na.omit(tit) 
```

```{r}
# fit a logistic regression with all variables without name and Home
# use glm()
fit_1.1 <- glm(Survived ~., data = tit.comp[,-c(1,10)], family = binomial(logit))
summary(fit_1.1)
```


```{r}
# fit a logistic regression with all variables without name, Fare, Parents and Home
# use glm()
fit_1.2 <- glm(Survived ~., data = tit.comp[,-c(1,7,8,10)], family = binomial(logit))
summary(fit_1.2)

```
```{r}
# Apply model selection by anova and Anova
anova_out <- anova(fit_1.1, fit_1.2, test = "Chisq")
Anova_out <- Anova(fit_1.1)
anova_out
```


```{r}
# conduct the final model
fit_1_final <- glm(formula = Survived ~ Passenger.Class + Sex + Age + Siblings.and.Spouses + 
    Port, family = binomial(logit), data = tit.comp[, -c(1, 10)])
summary(fit_1_final)
```
```{r warning = FALSE}
# Do prediction

# Create your jack

jack <- tit.comp[1,]
jack[1,] <- c("Jack","",3,"male",17,0,"","","S","")
jack$Age <- jack$Age %>% as.numeric()
jack$Siblings.and.Spouses <- jack$Siblings.and.Spouses %>% as.numeric()

# use predict()
predict(fit_1_final, jack, type = "response")

predict(fit_1_final, jack)

logodd = predict(fit_1_final, jack)
exp(logodd)/(exp(logodd)+1)
```
```{r, message=F}
# Draw the ROC curve for the final model

# roc_final <- roc(tit.comp$Survived, fit_1_final$fitted.values)

# # plot the roc curve with ggroc

# roc_final%>%
#   ggroc(colour = 'steelblue', size = 2) +
#   ggtitle(paste('ROC Curve ', '(AUC = ', round(roc_final$auc,4), ')'))
```



# Multinomial Logistic Regression


```{r}
# Y: n levels
## n-1 logistics regressions
## (n-1)*p parameters
```

## 
```{r}

# Use nnet to conduct Multinominal
# multinominal regression <=> single hidden layer neural network activated by sigmoid function

library(nnet)

```

## Nominal Response
```{r}
# Use WVS data from {carData}

fit_mult <- multinom(poverty ~ ., data = WVS)
str(WVS)
summary(fit_mult)
```

```{r}
Anova(fit_mult)
```

```{r warning = FALSE}
# Create a new sample
newdata <- WVS[1,]
newdata[1,] <- c("", "yes", "yes", "USA", 18 , "female")
newdata$age <- as.numeric(newdata$age)
```

```{r}
# Prediction
predict(fit_mult,newdata, type = "class")
predict(fit_mult,newdata, type = "prob")
```

## Ordinal Response
```{r}
pacman::p_load(MASS)
```

```{r}
# use polr to apply ordinal logistic
fit_ord <- polr(poverty ~ ., data = WVS)
summary(fit_ord)

Anova(fit_ord)

predict(fit_ord, newdata, type = "class")
predict(fit_ord, newdata, type = "prob")
```
















































