---
title: 'Case Study: Titanic Passengers'
author: 'MSBA7002: Business Statistics'
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
  html_document:
    code_folding: show
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE , fig.width=6, fig.height=4)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(bestglm, glmnet, leaps, car, tidyverse, mapproj, dplyr, GGally)
```

```{r include = FALSE}
# Help function, to draw a beautiful ggpairs() plot without `binwidth` warning
my_bin <- function(data, mapping, ..., low = "#132B43", high = "#56B1F7") {
  ggplot(data = data, mapping = mapping) +
    geom_bin2d(...) +
    scale_fill_gradient(low = low, high = high)
}
```

The sinking of the RMS Titanic is one of the most infamous shipwrecks in history. On April 15, 1912,
during her maiden voyage, the Titanic sank after colliding with an iceberg, killing 1,502 of the 2,224
passengers and crew. This sensational tragedy shocked the international community and motivated the
adoption of better maritime safety regulations.

One of the reasons that the shipwreck led to such loss of life was that there were not enough lifeboats for
the passengers and crew. Although there was some element of luck involved in surviving the sinking,
some groups of people were more likely to survive than others.

We use this rich and storied example to explore some questions of interest about Titanic survival rates.
For example, were there any key characteristics shared by survivors? Were some passenger groups
more likely to survive than others? Can we accurately predict survival? 

We will fit a logistic regression model using `titanicpassengers.dat` to explore these questions.

The data describes the survival status of 1,309 of the 1,324 individual passengers on the Titanic. Information on the 899 crew members is not included. The variables are described below: 

<center>
|Feature|Description|
|--|---|
|Name|Passenger Name|
|Survived|Yes or No|
|Passenger Class|1, 2, or 3 corresponding to 1st , 2nd , or 3rd  class| 
|Sex|Passenger gender|
|Age|Passenger age|
|Siblings and Spouses|The number of siblings and spouses aboard|
|Parents and Children|The number of parents and children aboard|
|Fare|The passenger fare|
|Port|Port of embarkment (C=Cherbourg; Q=Queenstown; S=Southampton)|
|Home/Destination|The home or final intended destination of the passenger|
</center>

\pagebreak
## 1. A quick EDA

Read `titanicpassengers-bbm.dat` as `tit`
```{r}
tit = read.csv('titanicpassengers-bbm.dat')
str(tit)
names(tit)
summary(tit)
```

Change the type of `Passenger.Class`, `Sex`, `Survived` from int into factor
```{r}
tit$Survived <- as.factor(tit$Survived)
tit$Passenger.Class <- as.factor(tit$Passenger.Class)
tit$Sex <- as.factor(tit$Sex)
str(tit)
```

Check the situation of missing value
```{r}
sum(is.na(tit))
sapply(tit, function(x) sum(is.na(x)))
```

A look at relationship between Passenger Class and Survived.
```{r}
plot(tit$Passenger.Class, tit$Survived, ylab ="Survived", xlab = "Passenger Class", col = c('blue','red'))
```

A look at relationship between Sex and Survived.
```{r}
plot(tit$Sex, tit$Survived, ylab ="Survived", xlab = "Sex", col = c('blue','red'))
```

We can also use 'ggpairs' to check relationships among the response and the predictors.
```{r, warning=FALSE}
tit[,c(2:7,9)] %>%
  ggpairs(progress=FALSE,lower=list(combo=wrap("facethist", binwidth=1), 
                                    continuous=wrap(my_bin, binwidth=0.25)))
```

\pagebreak
## 2. Multiple Logistic Regression

To avoid problems, we will delete all cases with missing values.
```{r}
tit.comp <- na.omit(tit)
```

### a. With all the predictors

#### i. Full Model
```{r}
fit1 <- glm(Survived~., tit.comp[,-c(1,10)], family=binomial(logit))
summary(fit1)
```

\pagebreak
#### ii. identify "problematic observations"

From the output, the Port variable appears strange. 

```{r}
table(tit.comp$Port)
```

From the output we notice two observations with empty entries for Fort, but they are not treated as missing values. So we need to replace them with NA. 

```{r}
tit.comp$Port=factor(tit.comp$Port, levels=c("C","Q","S"))
```

Refit the full model, and notice the different outputs about Port.

```{r}
fit1 <- glm(Survived~., tit.comp[,-c(1,10)], family=binomial(logit))
summary(fit1)
```

\pagebreak
To delete the insignificant predictors `Parents.and.Children` and `Fare`.
```{r}
fit0 <- glm(Survived~., tit.comp[,-c(1,7:8,10)], family=binomial(logit))
summary(fit0)
```
\pagebreak
#### iii. Are Parents.and.Children and Fare needed?

We can use the Likelihood Ratio test to see whether the variables are needed. The null hypothesis is that the smaller model is sufficient.

```{r}
anova(fit0,fit1,test="Chisq")
```

We can also calculate the Chi-square statistics directly.
```{r}
chi.sq.2 <- fit0$deviance-fit1$deviance
chi.sq.2
pchisq(chi.sq.2, 2, lower.tail=FALSE)
```

Both analyses suggest that we can use the smaller model. 

\pagebreak
### b. Final Model
```{r}
fit <- glm(Survived~., tit.comp[,-c(1,7:8,10)], family=binomial(logit))
summary(fit)
```

Practice how to interpret the estimated coefficients as odds ratio.

\pagebreak
### c. Prediction

Here we would like to predict the survival rate of Jack and Rose by using their information below:

<center>
|Name|Jack|Rose|
|--|---|---|
|Passenger Class|3|1|
|Sex|male|female|
|Age|17|20|
|Siblings and Spouses|0|1|
|Port|S|S|
</center>

```{r warning = FALSE}
jack <- tit.comp[1,]
jack[1,] <- c("Jack","",3,"male",17,0,"","","S","")
jack$Age <- jack$Age %>% as.numeric()
jack$Siblings.and.Spouses <- jack$Siblings.and.Spouses %>% as.numeric()
jack
```

```{r}
print(paste('Survival rate of Jack is',round(predict(fit,jack,type='response'),4)))
```


```{r warning = FALSE}
rose <- tit.comp[1,]
rose[1,] <- c("Rose","",1,"female",20,1,"","","S","")
rose$Age <- rose$Age %>% as.numeric()
rose$Siblings.and.Spouses <- rose$Siblings.and.Spouses %>% as.numeric()
rose
```

```{r}
print(paste('Survival rate of Rose is',round(predict(fit,rose,type='response'),4)))
```
